Positivity-Preserving Adaptive Runge--Kutta Methods

### 2020-04-08 nusslein_vsrp/latex/overview.tex
Pseudocode for algorithm

@@

At first, the stage values $y_i$ and stage derivatives $f_i$ are calculated.
Afterwards, $u^{n+1}$ is computed using the original weights $b$.
If the positivity is not violated, this solution is used to update the solution.
If $u^{n+1}$ contains negative values, the weights are adapted using the following algorithm.

Start with the highest order $\tilde p$ such that there are degrees
of freedom for the weights $\bt$. Apply an LP solver to the optimization
problem with accuracy and positivity constraints.
If the LP is infeasible, reduce the order $\tilde p$ by one.
If the LP is feasible, test whether the solution is acceptable.
At first, test whether the new $u^{n+1}$ satisfies the constraints.
This is necessary because sometimes the applied LP solvers incorrectly
identify the problem as feasible.
Afterwards the adaption error is calculated using \eqref{eq:Err}.
If the error is bigger than a given maximum error tolerance, the
new weights $\bt$ are also rejected.
If the new set of weights is rejected in one of the tests the order
$\tilde p$ is reduced by one.
As soon as the algorithm finds suitable weights, these are used to update the solution.
If all acceptable orders are tried and no weights were found, the step hast to be rejected.

@@

>>>>

%pseudocode
\usepackage[boxed]{algorithm}
\usepackage{algpseudocode}

@@

In this section an algorithm to update the solution $u^n$ using the adaptive RKM is proposed.
The pseudocode is given in Algorithm\,\ref{alg:Adaption}.
t
At first, the stage values $y_i$ and stage derivatives $f_i$ are calculated.
Afterwards, $u^{n+1}$ is computed using the original weights $b$.
If the positivity is not violated, this solution is used to update the solution.
If $u^{n+1}$ contains negative values, the weights are adapted using the following algorithm.

Start with $\tilde{p}$ equal to the highest order $p_{start}$ such that there are degrees of freedom for the weights $\bt$. 
Apply an LP solver to the optimization
problem with accuracy and positivity constraints.
If the LP is infeasible, reduce the order $\tilde p$ by one.
If the LP is feasible, test whether the solution is acceptable.
At first, test whether the new $u^{n+1}$ satisfies the constraints.
This is necessary because sometimes the applied LP solvers incorrectly
identify the problem as feasible.
Afterwards the adaption error is calculated using \eqref{eq:Err}.
If the error is bigger than a given maximum error tolerance, the
new weights $\bt$ are also rejected.
If the new set of weights is rejected in one of the tests the order
$\tilde p$ is reduced by one.
As soon as the algorithm finds suitable weights, these are used to update the solution.
If all acceptable orders are tried and no weights were found, the step hast to be rejected.

@@

\begin{algorithm}
\begin{algorithmic}[1]
\State Caclculate Stagedrivatives
\State $u^{n+1}=\Delta t Fb$
\If {$u^{n+1} \geq 0$} 
        \State Update and End
\Else
	\ForAll{$\tilde p \in [p_{start}:p_{min}]$} 
		\State Solve LP
		\If {LP Feasible}
			\State caclulate $err_{adapt}$
			\If {$err_{adapt} < tol$}
				\State $u^{n+1}=\Delta t F \tilde b$
				\State Update and End
			\EndIf 
		\EndIf
	\EndFor
	\State Reject step and End
\EndIf 
\end{algorithmic}
\caption{Adaption of weights}
\label{alg:Adaption}
\end{algorithm}
